Rhizaria in the oligotrophic ocean exhibit clear temporal and vertical variability.

### 2024-07-06 manuscript/02_methods.qmd
minor edit and render

The UVP5 samples at \~15Hz rate as it descends the water column and records the exact position of each particle larger than 600$\mu m$ (ESD). However, identified rhizaria ranged from a 934$\mu m$ ESD Aulacanthidae cell to a Collodarian colony over 10mm in diameter. To confirm that the UVP5 was sampling adequately across all size ranges, an normalized biomass size spectrum (NBSS) slope was constructed to identify a drop-off which would indicate poor-sampling at the small size range [@lombard2019; @barth2024]. However, it was evident from this analysis that all size ranges were adequately sampled across the size range (Supplemental Figure 2) so no data were excluded. The UVP5 reports the exact depth at which a particle is recorded, however to estimate abundance, observations must be binned over fixed depth intervals. Our deployments had variable descent depths and speeds with more casts descending to 500m than 1000m and descents quicker through the epipelagic than the mesopelagic (see @barth2022 for an extended discussion of UVP5 data processing). For the present study, Rhizaria abundances were estimated in 25m vertical bins, which offer a moderate sampling volume per bin (average 0.948$m^3$ in the epipelagic and 0.589$m^3$ in the mesopelagic; Supplemental Table 1) while still maintaining ecologically relevant widths. However, concentrations in a 25m bin would need to be greater than 2.428 ind. $m^{-3}$ and 3.912 ind. $m^{-3}$, in the epipelagic and mesopelagic respectively, to fall below a 10% non-detection risk [@benfield1996; @barth2024]. Because we typically observed many Rhizaria taxa below these concentrations, we present the 25m binned data to visualize broad-scale average distributions. For quantifying and modelling Rhizaria abundances, we present integrated abundance estimates, with each cast. Due to the variable descent depths of the UVP, data are categorized as epipelagic (0-200m), upper mesopelagic (200-500m), and lower mesopelagic (500-1000m). The average sampling volume integrated through these regions were 7.59$m^2$, 7.06$m^2$, and 11.77$m^2$, with non-detection thresholds at 0.30 ind. $m^{-2}$, 0.33 ind. $m^{-2}$, and 0.20 ind. $m^{-2}$ respectively. All UVP data processing was done using the `EcotaxaTools` package in R (Barth 2023).

>>>>

The UVP5 samples at \~15Hz rate as it descends the water column and records the exact position of each particle larger than 600$\mu m$ (ESD). However, identified rhizaria ranged from a 934$\mu m$ ESD Aulacanthidae cell to a Collodarian colony over 10mm in diameter. To confirm that the UVP5 was sampling adequately across all size ranges, an normalized biomass size spectrum (NBSS) slope was constructed to identify a drop-off which would indicate poor-sampling at the smaller end of the size range [@lombard2019; @barth2024]. However, it was evident from this analysis that all size ranges were adequately sampled across the size range (Supplemental Figure 2) so no data were excluded. The UVP5 reports the exact depth at which a particle is recorded, however to estimate abundance, observations must be binned over fixed depth intervals. Our deployments had variable descent depths and speeds with more casts descending to 500m than 1000m and descents quicker through the epipelagic than the mesopelagic (see @barth2022 for an extended discussion of UVP5 data processing). For the present study, Rhizaria abundances were estimated in 25m vertical bins, which offer a moderate sampling volume per bin (average 0.948$m^3$ in the epipelagic and 0.589$m^3$ in the mesopelagic; Supplemental Table 1) while still maintaining ecologically relevant widths. However, concentrations in a 25m bin would need to be greater than 2.428 ind. $m^{-3}$ and 3.912 ind. $m^{-3}$, in the epipelagic and mesopelagic respectively, to fall below a 10% non-detection risk [@benfield1996; @barth2024]. Because we typically observed many Rhizaria taxa below these concentrations, we present the 25m binned data to visualize broad-scale average distributions. For quantifying and modelling Rhizaria abundances, we present integrated abundance estimates, with each cast. Due to the variable descent depths of the UVP, data are categorized as epipelagic (0-200m), upper mesopelagic (200-500m), and lower mesopelagic (500-1000m). The average sampling volume integrated through these regions were 7.59$m^2$, 7.06$m^2$, and 11.77$m^2$, with non-detection thresholds at 0.30 ind. $m^{-2}$, 0.33 ind. $m^{-2}$, and 0.20 ind. $m^{-2}$ respectively. All UVP data processing was done using the `EcotaxaTools` package in R (<https://github.com/TheAlexBarth/EcotaxaTools?tab=readme-ov-file>).
